---
title: "statistical report TU Dortmund"
author: "Emil Stasevski"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



#### Loading libraries



```{r}
options(scipen = 999)
library(tidyverse)
library(stargazer)
library(psych)
library(car)
library(gridExtra)
```





### Data integrity

```{r}
df = read.csv("~/get out of Mordor/dortmund/passes.csv", sep = ';')

df %>% nrow()

df %>% head()
```


Checking NA values


```{r}
df %>% is.na() %>% colSums()
```

```{r}
df %>% 
  filter(is.na(passing_quote))
```

So basically, there are only two empty values, and, luckily only for only one team. 


```{r}
table(df$winner)
```








```{r}
df %>% 
  ggplot(aes(x = winner, fill = winner)) +
  geom_bar()+
  labs(y='Number of observations',x='Team winning status', 
       title='The proportion of won and lost teams')+
  theme_bw()+
   theme(axis.title.x = element_text(size=12, face='bold'),axis.title.y = element_text(size=12, face='bold'), title =element_text(size=14, face='bold') )+
  theme(axis.text.x = element_text( size=10))+
  theme(axis.text.y = element_text( size=10))

ggsave("~/get out of Mordor/dortmund/img/img1.jpg"  ,  dpi = 400
      )
df  %>% drop_na() -> df

```






```{r}
df %>% 
  ggplot(aes(x = passing_quote))+
  geom_histogram(fill = 'deeppink')+
  labs(y='Number of observations',x='Passing rate', 
       )+
  theme_bw()+
   theme(axis.title.x = element_text(size=12, face='bold'),axis.title.y = element_text(size=12, face='bold'), title =element_text(size=14, face='bold') )+
  theme(axis.text.x = element_text( size=10))+
  theme(axis.text.y = element_text( size=10))

ggsave("~/get out of Mordor/dortmund/img/img2.1.jpg",   dpi = 400)

df %>% 
  ggplot(aes(y = passing_quote))+
  geom_boxplot()+
  labs(x='', y ='Passing rate', 
       )+
  theme_bw()+
   theme(axis.title.x = element_text(size=12, face='bold'),axis.title.y = element_text(size=12, face='bold'), title =element_text(size=14, face='bold') )+
  theme(axis.text.x = element_text( size=10))+
  theme(axis.text.y = element_text( size=10)) ->p2


ggsave("~/get out of Mordor/dortmund/img/img2.2.jpg",  dpi = 400, width = 6, height = 8 )



```








```{r}
library(stargazer)


stats <- describeBy(df$passing_quote, df$winner, mat = T) 

stats 
```






```{r}
summary_by_status <- df %>% drop_na() %>% 
  group_by(winner) %>%
  summarize(
    mean_passing_quote = mean(passing_quote, na.rm = TRUE),
    sd_passing_quote = sd(passing_quote, na.rm = TRUE),
    min_passing_quote = min(passing_quote, na.rm = TRUE),
    max_passing_quote = max(passing_quote, na.rm = TRUE),
    n = n()
  )

# Convert summary data into a data frame
summary_long <- summary_by_status %>%
  pivot_longer(cols = -winner, names_to = "statistic", values_to = "value")

# Convert summary data into a data frame
summary_long_df <- as.data.frame(summary_long)

# Generate LaTeX table
stargazer(
  summary_long_df,
  type = "latex",
  summary = FALSE,
  title = "Summary Statistics by Status",
  label = "tab:summary_by_status",
  digits = 2,
  out = "summary_by_status.tex"
)
```





```{r}
df %>% 
  ggplot(aes(x = winner, y = passing_quote, fill = winner))+
  geom_boxplot()+
  labs(x='Team winning status',y ='Passing quote', 
       )+
  theme_bw()+
   theme(axis.title.x = element_text(size=12, face='bold'),axis.title.y = element_text(size=12, face='bold'), title =element_text(size=14, face='bold') )+
  theme(axis.text.x = element_text( size=10))+
  theme(axis.text.y = element_text( size=10)) 


ggsave("~/get out of Mordor/dortmund/img/img4.1.jpg",   dpi = 400)

df %>% 
  ggplot(aes( x = passing_quote, fill = winner))+
  geom_density()+
  labs(y='Density',x='Passing rate', 
       )+
  theme_bw()+
   theme(axis.title.x = element_text(size=12, face='bold'),axis.title.y = element_text(size=12, face='bold'), title =element_text(size=14, face='bold') )+
  theme(axis.text.x = element_text( size=10))+
  theme(axis.text.y = element_text( size=10)) 




ggsave("~/get out of Mordor/dortmund/img/img4.2.jpg",   dpi = 400)
```




```{r}
df %>% 
  ggplot(aes( x = passing_quote, fill = winner))+
  geom_density()+
  theme_bw()

```


```{r}
df %>% filter(passing_quote > 60) %>% nrow() /  df %>% nrow()
```




```{r}
leveneTest(passing_quote ~ winner, data = df)
```



```{r}
shapiro.test(df$passing_quote)
```

*So passing_quote in general, and its values for winning and loosing teams are NOT normally distributed.  We should apply not-parametric methods.*




```{r}
wilcox_test_result <- wilcox.test(passing_quote ~ winner, data = df)
wilcox_test_result
```




```{r}
kruskal_test_result <- kruskal.test(passing_quote ~ winner, data = df)
kruskal_test_result 
```





**Since the data is not normally distributed the further test's inference is low, is procured additionaly**

T-test


```{r}
t_test_result <- t.test(   passing_quote ~ winner , data = df)
t_test_result
```




Analysis of variance (ANOVA)


```{r}
m1 <- aov(passing_quote ~ winner, data = df)
summary(m1)
```








Is the expected difference in passing rate in games that end in a draw different from the expected difference in passing rate in games with a winner? 


```{r}
df %>% filter(winner == 'No') %>% select(game_id) %>% 
  filter(duplicated(game_id)) %>% as.vector() ->draw_games_id
```



```{r}
df$is_draw <- ifelse(df$game_id %in% draw_games_id$game_id, 'Yes, draw', 'No, one team won')
```




```{r}
table(df$is_draw)
```




```{r}
df %>% drop_na() %>% 
  group_by(is_draw) %>% summarise(n = n()) %>% 
  arrange(desc(is_draw)) %>%
  mutate(prop = n / sum(df$n) *100) %>%
  mutate(ypos = cumsum(prop)- 0.5*prop ) %>% 
  ggplot(aes(x = '', y = n, fill = is_draw))+
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
   geom_text(aes( y = ypos, label = n), color = "black", size=6) +
  theme_void()+
  labs(title = 'Game outcomes proportion', x = '', y = '', axis.text.y = '') +
  
  guides(fill=guide_legend(title="Is draw"))+
   theme(axis.title.x = element_text(size=12, face='bold'),axis.title.y = element_text(size=12, face='bold'), title =element_text(size=14, face='bold') )+
  theme(axis.text.x = element_text( size=10))+
  theme(axis.text.y = element_text( size=10),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()
        ) 



ggsave("~/get out of Mordor/dortmund/img/img3.0.jpg",   dpi = 400)
```


```{r}
summary_by_status <- df %>% drop_na() %>% 
  group_by(is_draw) %>%
  summarize(
    mean_passing_quote = mean(passing_quote, na.rm = TRUE),
    sd_passing_quote = sd(passing_quote, na.rm = TRUE),
    min_passing_quote = min(passing_quote, na.rm = TRUE),
    max_passing_quote = max(passing_quote, na.rm = TRUE),
    n = n()
  )

# Convert summary data into a data frame
summary_long <- summary_by_status %>%
  pivot_longer(cols = -is_draw, names_to = "statistic", values_to = "value")

# Convert summary data into a data frame
summary_long_df <- as.data.frame(summary_long)

# Generate LaTeX table
stargazer(
  summary_long_df,
  type = "latex",
  summary = FALSE,
  title = "Summary Statistics by Game outcome",
  label = "tab:summary_by_game_outcome",
  digits = 2,
  out = "summary_by_status.tex"
)
```







```{r}
wilcox_test_result <- wilcox.test(passing_quote ~ is_draw, data = df)
wilcox_test_result
```




```{r}
kruskal_test_result <- kruskal.test(passing_quote ~ is_draw, data = df)
kruskal_test_result 
```


```{r}
df %>% 
  ggplot(aes(x = is_draw, y = passing_quote, fill = is_draw))+
  geom_boxplot()+
  labs(x='Game result',y ='Passing quote'
       )+
  guides(fill=guide_legend(title="Is draw"))+
  theme_bw()+
   theme(axis.title.x = element_text(size=12, face='bold'),axis.title.y = element_text(size=12, face='bold'), title =element_text(size=14, face='bold') )+
  theme(axis.text.x = element_text( size=10))+
  theme(axis.text.y = element_text( size=10)) ->p2

ggsave("~/get out of Mordor/dortmund/img/img3.1.jpg",   dpi = 400)


df %>% 
  ggplot(aes( x = passing_quote, fill = is_draw))+
  geom_density()+
  labs(y='Density',x='Passing rate', 
       )+
  theme_bw()+
  guides(fill=guide_legend(title="Is draw"))+
   theme(axis.title.x = element_text(size=12, face='bold'),axis.title.y = element_text(size=12, face='bold'), title =element_text(size=14, face='bold') )+
  theme(axis.text.x = element_text( size=10))+
  theme(axis.text.y = element_text( size=10)) 



ggsave("~/get out of Mordor/dortmund/img/img3.2.jpg",   dpi = 400)
```



*Interesting games*

```{r}
df %>% filter(passing_quote < 60 & winner == 'Yes')
```





